---
title: "Plot dwgsim_eval output"
author: "Dave Tang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
    toc_float: false
---

```{r setup, include=FALSE}
library(tidyverse)
theme_set(theme_bw())
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = FALSE)
```

Create column vector based [on the documentation](https://github.com/nh13/DWGSIM/blob/main/docs/04_Evaluating_Mappings.md).

The `dwgsim_eval` tool outputs a table with the following columns:

| Number | Column | Description |
| ------ | ------ | ----------- |
|    1   |  `thr` | The alignment threshold (see `-a`) |
|    2   |  `mc`  | The number of reads mapped correctly that should be mapped at the threshold |
|    3   |  `mi`  | The number of reads mapped incorrectly that should be mapped be mapped at the threshold |
|    4   |  `mu`  | The number of reads unmapped that should be mapped be mapped at the threshold |
|    5   |  `um`  | The number of reads mapped that should be unmapped be mapped at the threshold |
|    6   |  `uu`  | The number of reads unmapped that should be unmapped be mapped at the threshold |
|    7   |  `mc + mi + mu + um + uu` | The total number of reads that should be unmapped be mapped at the threshold |
|    8   |  `mc'` | The number of reads mapped correctly that should be mapped at or greater than the threshold |
|    9   |  `mi'` | The number of reads mapped incorrectly that should be mapped be mapped at or greater than the threshold |
|   10   |  `mu'` | The number of reads unmapped that should be mapped be mapped at or greater than the threshold |
|   11   |  `um'` | The number of reads mapped that should be unmapped be mapped at or greater than the threshold |
|   12   |  `uu'` | The number of reads unmapped that should be unmapped be mapped at or greater than the threshold |
|   13   |  `mc' + mi' + mu' + um' + uu'` | The total number of reads that should be unmapped be mapped at or greater than the threshold |
|   14   |  `mc / (mc + mi + mu)` | Sensitivity at the threshold. I.e. the fraction of reads that should be mapped that are mapped correctly. |
|   15   |  `mc / (mc + mi)` | Positive predictive value at the threshold. I.e. The fraction of mapped reads that are mapped correctly. |
|   16   |  `um / (um + uu)` | False discovery rate at the threshold.  I.e. The fraction of random reads that are mapped. |
|   17   |  `mc' / (mc' + mi' + mu')` | Sensitivity at or greater than the threshold. I.e. the fraction of reads that should be mapped that are mapped correctly. |
|   18   |  `mc' / (mc' + mi')` | Positive predictive value at or greater than the threshold. I.e. The fraction of mapped reads that are mapped correctly. |
|   19   |  `um' / (um' + uu')` | False discovery rate at or greater than the threshold.  I.e. The fraction of random reads that are mapped. |

"At or greater than the threshold" tells us what our sensitivity, PPV, and FDR would be if we filtered based on that threshold.

```{r col_spec}
my_cols <- c(
  'mq', 'mc', 'mi', 'mu', 'um', 'uu', 'mt',
  'mct', 'mit', 'mut', 'umt', 'uut', 'mtt',
  'recall', 'ppv', 'fdr',
  'recallt', 'ppvt', 'fdrt'
)

my_col_types <- cols(
  'i', 'i', 'i', 'i', 'i', 'i', 'i',
  'i', 'i', 'i', 'i', 'i', 'i',
  'd', 'd', 'd',
  'd', 'd', 'd'
)
```

Load evaluation result.

```{r load_eval}
mm <- read_table("../test/hg38.mm.eval.txt.gz", comment = "#", col_names = my_cols, col_types = my_col_types)
bwa <- read_table("../test/hg38.bwa.eval.txt.gz", comment = "#", col_names = my_cols, col_types = my_col_types)
head(mm)
```

Join results.

```{r combine}
mm %>%
  inner_join(y = bwa, suffix = c(".mm", ".bwa"), by = "mq") %>%
  pivot_longer(cols = !mq, names_to = c(".value", "tool"), names_sep = "\\.") -> metrics
```

Accuracy.

```{r accuracy}
metrics %>%
  group_by(tool) %>%
  summarise(total_read = sum(mt), accuracy = sum(mc)/sum(mt))
```

Accuracy without mapping quality zero.

```{r accuracy_no_mapq_zero}
metrics %>%
  filter(mq > 0) %>%
  group_by(tool) %>%
  summarise(total_read = sum(mt), accuracy = sum(mc)/sum(mt))
```

Plot percentage mapped correctly at different mapping qualities.

```{r percent_correct}
ggplot(metrics, aes(mq, mc/mt*100, colour = tool)) +
  geom_point() +
  geom_line() +
  ylim(c(0, 100))
```

Sensitivity at or greater than the threshold, i.e. the fraction of reads that should be mapped that are mapped correctly.

```{r recallt}
ggplot(metrics, aes(mq, recallt, colour = tool)) +
  geom_point() +
  geom_line() +
  ylim(c(0.5, 1))
```

Positive predictive value at or greater than the threshold, i.e. The fraction of mapped reads that are mapped correctly.

```{r ppv}
ggplot(metrics, aes(mq, ppvt, colour = tool)) +
  geom_point() +
  geom_line()
```

False discovery rate at or greater than the threshold.  I.e. The fraction of random reads that are mapped. |

```{r fdrt}
ggplot(metrics, aes(mq, fdrt, colour = tool)) +
  geom_point() +
  geom_line()
```
